### 2013-02-11 doc/icml2013/algo.tex
Updated draft; intro, models, algo

\section{Algorithm}
\label{sec:algo}

Recovery of $B$, $\mathcal{B}$.

Use AHK2012

Lemma error of $B$.

Lemma AHK2012 error

Theorem on polynomial recovery

>>>>

\section{A Method of Moments Algorithm}
\label{sec:algo}

In this section, we will outline some shortcomings of the standard
method of moments formulation for the mixture of linear regressions
model. We then describe how we can work around this limitation and
reduce our problem to a known one, which can be solved using symmetric
tensor decomposition. Finally, we will state some implications this
reduction has regarding identifiability.

\subsection{Notation}

We have already defined $y$ to be the response variable and $x \in
\Re^d$ to be the covariates it depends on. Note that the matrix of
regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
$\beta_k$.

We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use $\odot$ to denote
the general dot product between $p$-th order tensors, i.e. $X \odot
Y = \sum_{i_1, i_2, \ldots i_p} X_{i_1, i_2, \ldots i_p} Y_{i_1, i_2,
\ldots i_p}.$ Similar to how $x^T A$ is the projection of a vector $x$
onto the row space of a matrix $A$, when $X \in \Re^{d^p}$ and $Y \in
\Re^{d^{p+q}}$,  with $p, q > 0$, $X \odot Y \in \Re^{d^q}$ is the
projection of $X$ onto the first $p$ dimensions of $Y$, i.e. $(X \odot
Y)_{i_1, \ldots, i_q} = \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p}
Y_{j_1, \ldots, j_p, i_1, \ldots, i_q}$.

The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.

\subsection{Identifiability}

The model is in general identifiable, except for a pathological case
concerning how data points are placed \citationneeded.

\subsection{Overview}

Conventionally, when using the method of moments, the moments of
observed variables are expressed algebraically in terms of the model
parameters. Using these equations, we attempt to solve for the model
parameters using empirical estimates of the moments.

In the case of the mixture of linear regressions, we have two observed
quantities, $x$ and $y$. Unfortunately, because $y$ is a scalar
quantity, we have only one equation for every moment of $y$, and with
this naive approach, would have to consider $O( (K+1) D)$ moments to
recover $B \in \Re{K \times D}$ and $\pi \in \Re^K$.

% To keep the discussion simple, we will consider the case where there is
% no noise, i.e.  $\sigma^2 = 0$. In this case, the moments of $y$ are,
% \begin{eqnarray*}
%   \E[ y ] &=& (\sum_{k=1}^{K} \pi_{k} \beta_k)^T x \\
%           &=& M_1^T \E[x] \\
%   \E[ y^2 ] &=& \sum_{k=1}^{K} \pi_{k} \E[(\beta_k^T x)]^2 \\
%   &=& \sum_{k=1}^{K} \pi_{k} (\beta_k\tp{2} \odot \E[x\tp{2}]) \\
%   &=& M_2 \odot \E[x\tp{2}] \\
%   \E[ y^p ] &=& M_p \odot \E[x\tp{p}].
% \end{eqnarray*}
% 
% We note that joint moments with $x$ will not help either; consider
% \begin{eqnarray}
%   \E[ y^p x^q ] &=& M_p \odot E[ x\tp{p+q} ].
% \end{eqnarray}
% The elements of $\E[ y^p x^q ]$ are linear combinations of $\E[y^p]$,
% and hence we do not get any new equations; \todo{This isn't exactly
% correct; should we even be presenting this line of reasoning?}. This is
% to be expected, since any information about the $\beta$s is really
% encoded in $y$, and not in the $X$.

To work around this problem, we will use regression on the conditional
moments, $\E[y | x]$ to recover the expected moments of the regressors,
$M_2$ and $M_3$. Recovering the $\beta_k$ and $\pi_k$ given these
symmetric moments has been addressed in prior work
\citep{AnandkumarHsuKakade2012, AnandkumarGeHsu2012}.  This algorithm as
been summarized in \algorithmref{algo:spectral-experts}.

\begin{algorithm}[t]
  \caption{Spectral Experts}
  \label{algo:spectral-experts}
  \begin{algorithmic}[1]
    \REQUIRE $\mathcal{D} = \{ (\xn{1}, \yn{1}), \cdots, (\xn{N}, \yn{N}) \}$
    \REQUIRE $K$, the number of clusters, $\sigma^2$ 
    \STATE Recover $M_1$ using least-squares regression on $\mathcal{D}$.
    \STATE Recover $M_2, M_3$ using low-rank regression on $\mathcal{D}$.
    \STATE Return the eigenvectors $\beta_k$ and eigenvalues $\pi_k$ of $M_3$ using $M_2$.
  \end{algorithmic}
\end{algorithm}

\subsection{Recovering $M_2$ and $M_3$}

Let $\epsilon \sim \normal{0, \sigma^2}$ be independent Gaussian noise.
we have,
\begin{eqnarray*}
  y &=& \sum_{k=1}^{K} \delta_{k} \beta_k^T x + \epsilon \\
  \E[ y | x ] &=& M_1^T x \\
  \E[ y^2 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^2 + \E[ \epsilon^2 ] \\
  &=& M_2 \odot x^{\otimes 2} + \sigma^2 \\
  \E[ y^3 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^3
            + 3 \E[\epsilon^2] (\sum_{k=1}^{K} \pi_{k} \beta_k^T) x \\
  &=& M_3 \odot x^{\otimes 3} + 3 \sigma^2 M_1.
\end{eqnarray*}

% Details about low rank regression.
The symmetric structure of $M_2$ and $M_3$ implies that there are
$\frac{d (d-1)}{2}$ and $\frac{d (d-1) (d-2)}{6}$ degrees of freedom
respectively. However, the moments are further constrained to have
a rank of $K$. To exploit this property, we will use trace norm
regularization with a proximal subgradient descent
algorithm\citationneeded to recover $M_2$,
\begin{eqnarray*}
  M_2 &=& \arg\min_{M} \half \sum_{i=1}^N ((\xn{i})\tp{2} \odot M - (\yn{i})^2)^2 + \lambda \|M\|_*.
\end{eqnarray*}

The analogue of trace norm regularization for higher order tensors
corresponds to the sum of the trace norms of the mode-k unfolding of the
tensor \cite{Tomioka2011}, $X_{(k)}$, is a $d_k \times (\prod_{k' \neq
k} d_{k'})$ matrix obtained by concatenating the entries for all
dimensions other than $k$. For example, the 1-mode unfolding of a 3rd
order tensor has entries, $X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2,
i_3}$. Thus, we employ a proximal subgradient descent algorithm to solve
the following optimization problem to recovery $M_3$.
\begin{eqnarray*}
  M_3 &=& \arg\min_{M} \half \sum_{i=1}^N ((\xn{i})\tp{3} \odot M - (\yn{i})^3)^2 
          + \frac{\lambda}{3} \sum_{k=1}^{3} \|M_{(k)}\|_*.
\end{eqnarray*}

In \sectionref{sec:theory}, we will show that recovery of matrices so
constrained requires only $O(\frac{\sqrt{K} d}{\epsilon})$ samples as
opposed to $O(\frac{d^3}{\epsilon})$ samples to recover moments within
$\epsilon$ from their true values.

\subsubsection{Implications for Identifiability}

Our approach employs regression on $x\tp{2}$ and $x\tp{3}$, and thus
requires $\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes
\mathcal{X} \otimes \mathcal{X}$ to be linearly independent modulo
symmetry. We call these condition {\em quadratic} and {\em cubic}
independence.

This condition trivially holds when the space $\mathcal{X}$ consists of
independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
places some constraints on non-linear featurizations. For example, the
common polynomial basis expansion, $1, x_1, x_1^2, \dots, x_1^p$ is
neither quadratically nor cubically independent because the product
space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$. In fact, the
number of unique terms in $x\tp{p}$ is only of the order of $p \times
d$, and hence we will not have enough equations to solve for $M_2$ or
$M_3$ unless we consider $p = d^2$ moments.

{\todo Note: we haven't shown identifiability from all moments. What
does this mean?}

\subsection{Recovering $B$ and $\pi$}

Having recovered the moments $M_2$ and $M_3$, we must now solve for the
actual regressors $\beta_k$ and mixture weights $\pi_k$, which are the
eigenvectors and eigenvalues of the symmetric moments respectively,
\begin{eqnarray*}
  M_2 &=&  \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 2}\\
  M_3 &=& \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 3}.
\end{eqnarray*}

We note that the eigendecomposition of $M_2$ is not sufficient to
recover the $\beta_k$ because they are not necessarily orthogonal to
each other. \citet{AnandkumarGeHsu2012} describes an approach that uses
rotates $M_3$ to an orthogonal basis by using the whitening transform of
$M_2$. The eigenvectors and eigenvalues recovered from the
eigendecomposition of $M_3(W, W, W)$ can be de-whitened to recover the
$\beta_k$ and $\pi_k$.

### 2013-02-11 doc/icml2013/intro.tex
rough changes
Updated draft; intro, models, algo

Local optima are a common problem plaguing the learning of parameters in
latent variable models. Recently, there has been an interest in spectral
algorithms that provide statistically consistent estimators, i.e. given
sufficient data they are guaranteed to converge to the global optimum,
given that the model is identifiable. Furthermore, these algorithms have
polynomial sample and computational complexities for several such
models, e.g.  hidden Markov models \cite{AnandkumarHsuKakade2012},
probabilistic context free grammars \cite{HsuKakadeLiang2012}, etc. In
this work, we provide a spectral algorithm for a discriminative model,
the mixture of linear regressions.

\todo{Where are mixtures of experts used? Why are they interesting?}

Our approach uses regression to recover the moments of the data from the
scalar observable output variable. Once we recover the second and third
moments of the regression coefficients, we use a slight modification of
the method outlined in \citet{AnandkumarHsuKakade2012}.

We evaluate our algorithm on simulated linear and non-linear data to
understand how well the algorithm scales with the number of components,
dimensions and separation. To test how the algorithms perform on a more
real dataset, we study their performance on a motion tracking dataset.

In \sectionref{sec:model}, we formally define the mixture of linear
regressions model and discuss when it is identifiable. We describe our
algorithm and provide some theoretical results in \sectionref{sec:algo}. We
present the results of our evaluations in \sectionref{sec:evaluation}.
Finally, we discuss the merits and demerits of spectral algorithms in
discriminative models and conclude in \sectionref{sec:discussion}. The
details of some proofs can be found in \sectionref{sec:proofs}.

>>>>

% Latent variables are important
% Local optima is problem
% Recent method of moments spectral methods sidestep local optima

Latent variable models have routinely been used in probabilistic
modelling, (examples). Parameters for these models have been learnt
through a variety of methods, EM, variational methods, etc. The local
nature of these algorithms lead them to converge to local optima.
Recently, there has been an interest in moment of methods estimators
with polynomial sample and computational complexities. These estimators
are statistically consistent, i.e. given sufficient data they are
guaranteed to converge to the global optimum in an identifiable model,
thus sidestepping the local optima issue altogether.

% Discriminative models work better in practice
% Discriminative models with latent variables add expressive power (examples)
This approach has been applied to a variety of generative models, e.g.
hidden Markov models \cite{AnandkumarHsuKakade2012}, probabilistic
context free grammars \cite{HsuKakadeLiang2012}, etc. However, in
practice it has been observed that discriminative models work better in
terms of parameter recovery\citationneeded, particularly in the large
sample regime. \todo{Examples! CRFs, etc.}

% This paper: first step in establishing provably correct estimation
% with simple mixture of linear regression
In this work, we provide a first step in establishing provably correct
estimation for such discriminative models by proposing an algorithm for
the simple mixture of linear regressions. The mixture of linear
regressions model \todo{models ?}. We describe the model formally in more
detail in \sectionref{sec:model}.

% Previous work: moment equations solvable by eigendecomposition or
% tensor factorization
While method of moment estimators were \todo{described by Fisher(?)},
they required as many moments as there were parameters, and hence deemed
impractical. Recently, \citet{AnandkumarHsuKakade2012} showed how using
eigendecomposition, the moment equations could be solved using just the
first three moments of data.

% Challenge: E[X^m Y^n] doesn't work
In the discriminative setting, we are trying to learn the parameters for
a model for $y | X$. However, as the response variables are just
scalars, it is not possible to recover the moments simply from $\E[ X^m
y^n ].$ At the same time, knowing the $X$ should give us more
information than the scalar $y$.

% Approach: Derive moment functions based on linear regression on Y^2
% and Y^3 Algorithm: trace norm regularization to recover tensor
% + tensor factorization
We propose an algorithm in \sectionref{sec:algo} that derives moment
functions based on linear regression on the moments of $y$, i.e.
$\E[y^2]$ and $\E[y^3]$. We exploit the low rank properties of the
second and third order moments to efficiently recover parameters, and
then recover the individual regressors, $\beta_k$, using the tensor
factorization method \cite{AnandkumarGeHsu2012}.

% Show some identifiability results; require quadratic and cubic
% independence
We prove that this approach is indeed statistically consistent with
polynomial sample complexity in \sectionref{sec:theory}. Aside from
standard identifiability results \citationneeded, we also show that
recovery using moments places some independence requirements on the
variables $X$.

% Empirical evaluation
% It works
%  EM is sensitive to initialization (local procdure)
%  Spectral provides good initialization
% Intuition: basin of attraction, local methods like EM improve

We evaluate our algorithm on simulated linear and non-linear data to
understand how well the algorithm scales with the number of components,
dimensions and separation. To show how robust the algorithm is to model
misspecification, we evaluate it's performance on \todo{a motion
tracking dataset(?)}.

% What properties of the problem make it hard on EM versus spectral?
%  Separation, dimensionality, number of clusters, etc.
% Real dataset?

We compare our approach with expectation-maximization; our results,
presented in \sectionref{sec:evaluation} shows that spectral methods do
indeed provide an answer close enough to the global optima that using
a local method from there recovers the true parameters. \todo{Describe
some of our other high-level findings.}

\todo{What about talking about initialization? Generalizing and kmeans++}

### 2013-02-11 doc/icml2013/model.tex
Updated draft; intro, models, algo

Mixture of Experts

\subsection{Identifiability}

>>>>

\newcommand{\xn}[1]{x^{(#1)}}
\newcommand{\xni}{\xn{i}}
\newcommand{\yn}[1]{y^{(#1)}}
\newcommand{\yni}{\yn{i}}

The mixture of linear regressions model describes response variables $y
\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
modes which occur with probability $\pi_1, \dots, \pi_K$ respectively.
Each mode is independently described by regression coefficients
$\beta_k$ and we consider a common variance term $\sigma^2$. Given $x$,
the generative process for the response variables is,
\begin{eqnarray*}
  h &\sim& \mult(\pi) \\
  y | h = k &\sim& \normal{ \beta_{k}^T x}{ \sigma^2 },
\end{eqnarray*}
where $h$ is a latent variable corresponding to the mode.

Our learning framework can be described as follows; we observe $N$
samples, $(\xn{1}, \yn{1}), \dots, (\xn{N}, \yn{N})$ and would like to
recover the regressors, $B = [\beta_1 \mid \beta_2 \mid \cdots \mid
\beta_K]$, and the mixture probabilities, $\pi = (\pi_1, \pi_2, \dots,
\pi_K)$. Note that we do not observe $h$. If we did, the optimal
solution for $\beta_k$ would be the regressor found using least squares
on the set of points, $\{(x_i, y_i) | h_i = k \}$.
